---
title: "Day 9 Session"
format: html
editor: visual
execute:
  warning: FALSE
---

## Session 2

```{r}
library(kableExtra)
library(patchwork)
library(tidyverse)
library(palmerpenguins)
library(reprex)
```

### Warm up

Starting from penguins create a new subset in which u only keep columns (species, island, sex, body mass, flipper length

filter to only include chinstrap and gentoo

rename the species column to "penguin_species"

Create a summary table that contains the mean of body mass, the mean flipper length, and the total count of observations each by specific spsecies

```{r}
penguin_col <- penguins %>% 
  select(c("species","island", "sex", "flipper_length_mm", "body_mass_g", ))

penguin_spec <- penguins %>% 
  filter(species %in% c("Chinstrap", "Gentoo")) %>% 
  rename(penguin_species = species) %>% 
  group_by(penguin_species,sex) %>% 
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE),
            mean_length = mean(flipper_length_mm, na.rm = TRUE),
            sample_size = n())

penguin_summm <- penguins %>% 
group_by(species,sex) %>% 
summarize(mean_body_mass = mean(body_mass_g, na.rm = TRUE),
mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            sample_size = n())


```

## Across from dplyr

Create a summary table of mean values for all numberic tables in he penguins data set

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```

Calculate the mean value of all columns hat start with bill

```{r}
penguins %>% 
  group_by(species, island) %>% 
  summarize(across(starts_with("bill"), mean, na.rm = TRUE))
```

write code that will return the min value pf any columns that end with mm grouped by island and year

```{r}
penguins %>% 
  group_by(island,year) %>% 
  summarize(across(ends_with("mm"), min, na.rm = TRUE))
```

Find and return both the mean and sd of all columns starting with bill grouped by penguin species and returning a new name that combines the original column with the function

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(where(is.numeric),
                   list(mean = mean, stddev = sd), 
                   na.rm = TRUE, 
                   .names = "{.fn}_{.col}"))
```

```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(starts_with("bill"),list(mean = mean, std_dev = sd), na.rm = TRUE, .names = "{.col}_{.fn}"))
```

```{r}
penguins_table <- penguins %>% 
  group_by(species, sex) %>% 
  summarize(mean_mass = mean(body_mass_g, na.rm = TRUE),
            sd_mass = sd(body_mass_g, na.rm  = TRUE)) %>% 
  kable(col.names = c("Species", "Sex", "Mean Body Mass (g)", "Standard Deviation of Body Mass (g)"))

penguins_table
```

```{r}
chart1 <- penguins %>% 
  ggplot(aes(x = body_mass_g, y = flipper_length_mm))+
  geom_point()
```

```{r}
chart2 <- penguins %>% 
  ggplot(aes(x = bill_length_mm))+
  geom_histogram()

```

```{r}
(chart1+chart2)/chart2
```

```{r}
#| fig-cap: Penguin body mass observatios by species
#| fig-alt: Jitterplot showing body mass measurements for three penguin species obsereved near palmer insert trands here
ggplot(penguins, aes(x = species, y = body_mass_g))+
  geom_jitter(width = 0.2, alpha = 0.5, color = "blue")
```
